---
title: "Lab 4"
author: "Ravi Patel, rspatel"
date: "March 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
offensive <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
biograpghy <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
```

Question 1:
Cleaning the data.
```{r}
defense$Name <- as.factor(defense$Name)
defense$Opponent_Opponent <- as.factor(defense$Opponent_Opponent)

offensive$Name <- as.factor(offensive$Name)
offensive$Opponent_Opponent <- as.factor(offensive$Opponent_Opponent)

biograpghy$Name <- as.factor(biograpghy$Name)
```

Question 2:
Setting the correct types.
```{r}

defense <- defense %>%
  mutate_if(is.character, as.numeric)

offensive <- offensive %>%
  mutate_if(is.character, as.numeric)

```

Creating new df variables according to the lab document and showing each's structure.
```{r}
defClean <- defense
offClean <- offensive
bioClean <- biograpghy

str(defClean)
str(offClean)
str(bioClean)
```

Tidying the data

Question 1:
```{r}
defClean <- gather(defClean, "Statistic", "Value", -c(1,2))
```

Quesiton 2:
Based upon the graphs, it is rare to have more than one count of the three turnover types.
```{r}
library(ggplot2)
ggplot(defClean, aes(x=Value)) +
  geom_histogram(binwidth = 1) +
  facet_wrap(~Statistic)
```

Question 3:
It appears that ISU played better defense versus Kansas State.
```{r}
def3 <- defClean %>%
  filter(Statistic == "Tackles_Solo") %>%
  filter(Opponent_Opponent == "West Virginia" | Opponent_Opponent == "Kansas State") %>%
  spread(Opponent_Opponent, Value)

def3[is.na(def3)] <- 0

ggplot(def3, aes(x=`West Virginia`, y=`Kansas State`)) + 
  geom_count() +
  labs(
    x = "West Virgina",
    y = "Kansas State",
    title = "Tackles Solo per Player per University"
  )
```

Tidying Part 2

Question 1:
```{r}
bio <- separate(bioClean, Hometown, c("City", "State"), sep=",")
head(bio)
```

Question 2:
```{r}
PlayersPerState <- bio %>%
  group_by(State) %>%
  summarise(
    n = n()
  )

ggplot(PlayersPerState, aes(x=State, y=n)) +
  geom_bar(stat="identity") + 
  labs(
    x = "State",
    y = "Number of Players",
    title = "Number of Player per State"
  ) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


